Evidence of self-care tooling on digital video platforms and phylogenetic modelling reveal parrot tool use is not rare

### 2024-08-19 manuscript.Rmd
revise manuscript

---
title             : "Digital video platforms and phylogenetic modelling reveal parrot tool use is not rare"
shorttitle        : "Parrot tool use"

@@

abstract: |
 Putatively rare behaviours, such as tool use, are challenging to study because absence of evidence can arise either from a species’ inability to produce the behaviour or from insufficient research effort. Here, we tackle this challenge by combining data from digital platforms and phylogenetic modelling to estimate actual rates of tool use in parrots. Videos on a social media platform revealed novel instances of tool use in 17 parrot species, more than doubling the confirmed number of tool-using parrot species from 11 (3%) to 28 (7%). Phylogenetic modelling ranked additional species that are most likely to be unobserved tool users, suggesting that between 11% and 17% of extant parrot species may be tool users. These discoveries have implications for inferences about the evolutionary drivers and origins of tool use in parrots, revealing associations with relative brain size and feeding generalism and indicating several genera where tool use was likely an ancestral trait. Overall, our findings challenge the assumption that current sampling effort captures the full distribution of putatively rare animal behaviours. Combining our sampling and analysis methods offers a fruitful approach for investigating the distribution, drivers, and origins of other rare behaviours.

@@

Here, we apply these approaches to tool use in the parrot order. We focus on 
tool use in parrots for a number of reasons. First, the scientific literature 
suggests that only a small proportion of extant parrot species (11 out of 398;
3%) use tools [@Auersperg2009;@Auersperg2010;@Auersperg2011a;@Auersperg2011b;
@Auersperg2012;@Bastos2021;@BentleyCondit2010;@Borsari2005;@Goodman2018;
@Heinsohn2017;@Janzen1976;@Lambert2015;@Lefebvre2002;@OHara2021]. Parrot tool 
use thus provides an ideal test case for examining how robust sampling is in the 
scientific literature. Second, parrots are highly popular as pets. Over 70% of 
all extant parrot species are bred in the aviculture industry and kept as pets 
worldwide [@Anderson2003;@Carrete2008;@Drews2001;@Kelly2014;@Li2014;@Su2015;
@Young2012], enabling us to leverage the power of digital video platforms to 
search for evidence of tool use [@Nelson2013]. Third, detailed data on relative
brain sizes [@Flammer2001;@Iwaniuk2005;@Mazengenya2018;@Olkowicz2016;
@Schuck2008;@Silva2017], feeding strategies [@Wilman2016], and shared ancestry
[@Jetz2012] exist for parrots, allowing us to fit the statistical model implied
by Figure \@ref(fig:plotDAG) to the entire parrot order.

We first present the results from our video survey, in which we collated videos
of tool use in parrots from a digital video platform. This survey reveals a 
number of previously unidentified tool-using parrot species, which we map onto 
the phylogeny of the parrot order. We then describe our statistical model in 
more detail, and use it to (_i_) rank further parrot species that are likely 
unobserved tool users and (_ii_) re-examine key hypotheses regarding the 
evolutionary drivers and origins of tool use in parrots.

@@

## Digital video platform reveals tool use in additional parrot species

@@

In total, we found `r nrow(crowdData)` videos of 
`r crowdData %>% pull(IndividualID) %>% unique() %>% length()` individuals from 
`r crowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
parrot species performing behaviours that met the definition of either true tool 
use (`r nrow(trueCrowdData)` videos of 
`r trueCrowdData %>% pull(IndividualID) %>% unique() %>% length()` 
individuals from 
`r trueCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
species) or borderline tool use (`r nrow(bordCrowdData)` videos of 
`r bordCrowdData %>% pull(IndividualID) %>% unique() %>% length()` 
individuals from 
`r bordCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
species). All videos featured pet parrots in captive settings. In 
`r crowdData %>% pull(HumanInfluence) %>% str_starts("No") %>% sum()` of these 
videos, owners did not appear to interact with the subjects. In 
`r crowdData %>% pull(HumanInfluence) %>% str_starts("Yes") %>% sum()` videos, 
there was potential human interaction, either from the owners being in close physical 
contact with the bird (e.g., bird perching on hand), talking to the bird, or 
handing it the tool (which occurred in only two videos). We could not establish 
the degree of human interaction in the remaining 
`r crowdData %>% pull(HumanInfluence) %>% str_starts("Unknown") %>% sum()` 
videos, as sound had been removed or was substituted by music. None of the 
videos featured owners directly rewarding tool use behaviours with food. All 
borderline tool use cases were excluded from further analyses.

Of the `r trueCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
parrot species performing true tool use, 
`r trueCrowdData %>% filter(Species != "Unknown") %>% group_by(Species, IndividualID) %>% summarise(n = n(), .groups = "drop") %>% group_by(Species) %>% summarise(check = n() > 1, .groups = "drop") %>% pull(check) %>% sum()` 
were represented in our video survey by two or more individuals over multiple 
independent observations. True tool use always involved the subject using an 
object for self-scratching (`r trueCrowdData %>% mutate(BodyPartScratched = str_detect(BodyPartScratched, "Head/Neck")) %>% pull(BodyPartScratched) %>% sum()` 
videos involved scratching the head and/or neck). The most common tool (`r sum(pull(trueCrowdData, EvaluatedTool) == "Feather")`
videos) was a moulted feather. Human-made objects (e.g., pens, spoons, pieces of 
wood, cardboard) were also common. In `r sum(readxl::read_xlsx(tar_read(crowdsourcing_file))$Directionality == "Tool to body", na.rm = TRUE)` 
videos, parrots manipulated the tool while keeping their body still, rather than 
holding the tool still and moving towards it.

According to YouTube video descriptions and owner comments, 
`r trueCrowdData %>% group_by(IndividualID, Sex) %>% summarise(.groups = "drop") %>% pull(Sex) %>% (function(x) sum(x == "Male"))` 
of the individuals performing true tool use were males and 
`r trueCrowdData %>% group_by(IndividualID, Sex) %>% summarise(.groups = "drop") %>% pull(Sex) %>% (function(x) sum(x == "Female"))` 
were females. No sex information was provided for the remaining 
`r trueCrowdData %>% group_by(IndividualID, Sex) %>% summarise(.groups = "drop") %>% pull(Sex) %>% (function(x) sum(x == "Unknown"))` 
individuals. As owners provided no information on whether sex had been 
established through genetic testing, and sexual dimorphism in parrots is 
rare [@Bercovitz1987;@Hoyo2011], we could not typically ascertain if descriptions 
were accurate. It is unclear if the disproportionately large number of males in 
the sample is a consequence of owners more readily assuming their parrots are 
male when they have not been genetically tested, owners being more likely to own 
or film male parrots, or male parrots exhibiting more true tool use behaviours 
than female parrots.

Figure \@ref(fig:plotPhylo1) maps the findings from the video survey onto a 
maximum clade credibility phylogeny for the parrot order, plotted alongside 
species previously identified in the scientific literature. Before the video 
survey, 11 parrot species (3%) had been identified as tool users in the 
scientific literature. Across our video survey, we observed true tool use in 
`r trueCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
species, 5 of which overlapped with the scientific literature and 
`r (trueCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()) - 5` 
of which were novel species. All of the species identified in the video survey 
were cockatoos (_Cacatuidae_), Old World parrots (_Psittacinae_), or neotropical 
parrots (_Arinae_). The most common species in our survey, accounting for 
`r crowdData %>% filter(Species == "Pyrrhura molinae") %>% nrow()` videos from 
`r crowdData %>% filter(Species == "Pyrrhura molinae") %>% pull(IndividualID) %>% unique() %>% length()` 
individuals, was the green-cheeked conure (_Pyrrhura molinae_). In accordance 
with the scientific literature, the video survey found no evidence of tool use 
in any species of Psittaculidae, despite this family containing some of the most 
commonly kept pet species, including lovebirds, lorikeets, and Asian parakeet 
species. Combining both the video survey and the scientific literature, we can 
thus identify 28 tool-using parrot species overall (7%), compared to the 11 
previously reported.

@@

The identification of new tool-using species in our video survey increases the 
extent to which phylogeny can explain the distribution of tool use in the parrot 
order. We estimated phylogenetic signal (Pagel's $\lambda$) of tool use using 
both the pre-video-survey and post-video-survey data. Pagel's $\lambda$ varies 
between 0 and 1, where 0 implies that the distribution of a trait across species 
is unexplained by phylogenetic relatedness and 1 implies that the distribution 
of a trait across species is fully explained by phylogeny. Using the evidence of 
tool use from the scientific literature alone (pre-video-survey data; 11 
tool-using species), we estimated an average posterior Pagel's $\lambda$ of 
`r printnum(mean(tar_read(phySignal1)))` (95% credible interval [`r printnum(quantile(tar_read(phySignal1), 0.025))` 
`r printnum(quantile(tar_read(phySignal1), 0.975))`]; total _n_ = 
`r nrow(tar_read(d))` species). This estimate was moderate-to-strong, but highly 
uncertain. In comparison, combining the evidence from both the literature and 
the video survey (post-video-survey data; 28 tool-using species) resulted in a 
stronger and more certain estimate of phylogenetic signal. With these data, we 
estimated Pagel's $\lambda$ = `r printnum(mean(tar_read(phySignal2)))` (95% CI 
[`r printnum(quantile(tar_read(phySignal2), 0.025))` 
`r printnum(quantile(tar_read(phySignal2), 0.975))`]; total _n_ = 
`r nrow(tar_read(d))` species). Thus, the results of our video survey increase 
the extent to which the distribution of tool use across parrot species can be 
explained by shared phylogenetic ancestry. This suggests that we can potentially 
use phylogenetic information, along with other variables, to identify further 
tool-using parrot species that may remain undetected.

@@

In addition to the 28 tool-using species identified in the literature and our 
video survey, we fitted a Bayesian phylogenetic survival cure model to rank 
further species that are likely to be undetected tool-users (i.e., tool-using 
species with no tool use reported in the literature or in videos).

@@

Our tool use problem has the same features. We are modelling a time-to-event; 
specifically, the amount of "time" (i.e., observation opportunities measured as 
the number of published papers or videos) until tool use is identified. This is
right-censored data because many species will not have had tool use identified
when we measure them. Moreover, we can assume that a certain proportion of the
population will never experience the event -- that is, they are a non-tool-user
and so we will never identify tool use no matter how long we measure them for.

@@

The posterior probabilities shown in Figure \@ref(fig:plotSurvCure1) are 
estimated with uncertainty, so it is difficult to "identify" any particular 
species as an undetected tool user. Nevertheless, taking the sum of all the 
posterior probabilities for the 
`r nrow(plotSurvCure1$data %>% filter(ToolUse == "Absent"))` species without 
observed evidence of tool use, we can estimate that around 
`r printnum(median(sumProbs), digits = 0)` of those species are likely to be 
undetected tool users (median sum of probabilities = 
`r printnum(median(sumProbs), digits = 2)`, 95% CI
[`r printnum(quantile(sumProbs, 0.025), digits = 2)` 
`r printnum(quantile(sumProbs, 0.975), digits = 2)`]). When combined with the 
species known to use tools, this implies that between 11% and 17% of extant 
parrot species may be tool users.

@@

Since the earliest anecdotes of parrots using tools by Wallace in the 1880s
[@Wallace1869] and more systematic anecdotal reports in the 1970s [@Goodall1971], 
only 11 parrot species (3% of all extant parrots) have been documented as tool 
users in the scientific literature. Our study collated data from a digital video
platform to identify 17 additional tool-using parrot species that are new to 
science, more than doubling the overall count to 28 species (7%). These species 
consisted of cockatoos (_Cacatuidae_), Old World parrots (_Psittacinae_), and 
neotropical parrots (_Arinae_). Beyond the video survey, the strong 
phylogenetic signal in our dataset allowed us to use phylogenetic information, 
along with other variables, to infer the unobserved probabilities of tool use 
across the parrot order. Our phylogenetic survival cure model incorporated
information on phylogenetic history, research effort, relative brain size, and 
feeding specialisation to rank parrot species that were most likely to be 
undetected tool users. The sum of probabilities from this model implied that 
between 15 and 41 of the species without observed evidence of tool use are 
likely to be undetected tool users, suggesting that the true proportion of tool
users may be as high as 17%.

These findings have a number of important implications. First, our findings show 
that current research effort in the scientific literature is insufficient to 
capture the real world occurrence of parrot tool use. If the scientific 
literature had sampled the natural world sufficiently, we would expect to see 
close correspondence between those species reported as tool users in the 
literature and those species the public have filmed performing tool use. 
Instead, we discovered a large discrepancy between these two data sources, both 
in the prevalence of tool use and the species identified, likely due to the 
numerous difficulties inherent to conducting observations in the wild. This 
raises the possibility that other rare behaviours remain yet to be discovered in 
the scientific literature.

@@

All instances of tool use in our video survey met the established criteria for 
tool use in the literature: the manipulation of an unattached object as an 
extension of the animal's body to achieve a goal [@Shumaker2011]. However, a 
striking feature of the dataset is that the 17 additional species identified in
the video survey were exclusively observed using tools for self-scratching.
While parrots in the wild use tools to achieve a variety of goals, including 
extractive foraging and preening [@Goodman2018;@OHara2021], parrots in captivity
only require tools to achieve the latter goal. It could be argued that 
self-scratching is a more cognitively simplistic form of tool use than 
extractive foraging as it is egocentric, meaning that the tool is directed 
towards oneself rather than an external object. Nevertheless, self-scratching 
meets the definition for a more complex type of embodied tool use known as 
'tooling', the deliberate generation of a mechanical interface by using an 
object to manipulate another target or surface [@Fragaszy2018]. In our dataset, 
most self-scratching tooling instances involved manipulating the tool while 
keeping the body still, suggesting that these were goal-directed, deliberate 
movements by individuals. This makes it difficult to dismiss self-scratching as 
unimportant: to do so would require a redefinition of tool-use and tooling in 
non-human animals. Moreover, the fact that we found strong phylogenetic 
clustering of self-scratching tool use from the video survey with other examples 
of tool use from the literature supports a common underlying cognitive 
mechanism. In line with this, some of the species of parrots that use tools for 
self-scratching in captivity also use tools for other purposes in the wild 
[@Auersperg2012;@Bastos2021;@Goodman2018;@OHara2021].

@@

Our data reveals that a wide range of parrot species exhibit self-scratching 
tool use behaviours in captive environments. It is possible that these self-care 
behaviours emerge more often in captivity because pet parrots are kept with few 
or no conspecifics, and therefore must innovate tool use to preen inaccessible 
body areas, such as their heads and necks, which would have been otherwise 
preened by flock members. This may indicate that human interaction and 
allopreening in the absence of a conspecific is insufficient for pet parrots, as 
suggested by studies on welfare and chronic stress in lone captive parrots 
[@Aydinonat2014;@Engebretson2006;@Mason2010;@Meehan2003]. Beyond social 
isolation in captivity, other factors that reduce preening efficacy by an 
individual might also lead to the innovation of self-care tooling: for example,
a captive kea housed with conspecifics in a naturalistic environment was shown
to have innovated the use of a pebble tool for self-care likely as a consequence 
of a missing upper mandible [@Bastos2021]. These behaviours might therefore be 
innovated by individuals when ecologically necessary. Critically, however, a 
cognitive capacity to innovate a behaviour such as tool use must exist in a 
species in order to manifest in any environment, whether that be captive or in 
the wild. As such, despite differences in individuals' life experiences in 
captive and wild settings, such behaviour emerging in a captive setting at the 
very least indicates an underlying cognitive capacity to innovate that same 
behaviour in the wild.

@@

In conclusion, we have shown that the scientific literature has insufficiently 
captured the full distribution of tool use in the parrot order. Our 
digital video survey more than doubled the number of known tool using parrot 
species from 11 to 28, and our phylogenetic model suggested that the true 
proportion of parrot tool users could be as high as 17% of all species in this 
order. These discoveries have implications for theories of the evolutionary 
drivers and origins of tool use in parrots. Beyond parrot tool use, the 
methods used in this study have the potential to be applied to other rarely 
observed behaviours, including tool use in other taxa [@Hunt2013], rhythmic 
entrainment in birds [@Benichov2016;@Patel2009;@tenCate2016;@Wilson2016],
teasing behaviours in primates [@Eckert2020], and tactical deception across all
animals [@BroJorgensen2010;@Byrne1988;@Byrne1991]. We hope that these methods
will continue to uncover a diverse array of ephemeral behaviours that have as
yet gone undetected in the scientific literature.

>>>>

---
title             : "Evidence of self-care tooling on digital video platforms and phylogenetic modelling reveal parrot tool use is not rare"
shorttitle        : "Parrot tool use"

@@

abstract: |
 Putatively rare behaviours, such as tool use, are challenging to study because absence of evidence can arise either from a species’ inability to produce the behaviour or from insufficient research effort. Here, we tackle this challenge by combining data from digital platforms and phylogenetic modelling to estimate actual rates of tool use in parrots. Videos on a social media platform revealed novel instances of self-care tooling in 17 parrot species, more than doubling the confirmed number of tool-using parrot species from 11 (3%) to 28 (7%). Phylogenetic modelling ranked additional species that are most likely to be unobserved tool users, suggesting that between 11% and 17% of extant parrot species may be capable of tool use. These discoveries have implications for inferences about the evolutionary drivers and origins of tool use in parrots, revealing associations with relative brain size and feeding generalism and indicating several genera where tool use was likely an ancestral trait. Overall, our findings challenge the assumption that current sampling effort captures the full distribution of putatively rare animal behaviours. Combining our sampling and analysis methods offers a fruitful approach for investigating the distribution, drivers, and origins of other rare behaviours.

@@

Here, we apply these approaches to tool use in the parrot order. We focus on 
tool use in parrots for a number of reasons. First, the scientific literature 
suggests that only a small proportion of extant parrot species (11 out of 398;
3%) use tools [@Auersperg2009;@Auersperg2010;@Auersperg2011a;@Auersperg2011b;
@Auersperg2012;@Bastos2021;@BentleyCondit2010;@Borsari2005;@Goodman2018;
@Heinsohn2017;@Janzen1976;@Lambert2015;@Lefebvre2002;@OHara2021]. Parrot tool 
use thus provides an ideal test case for examining how robust sampling is in the 
scientific literature. Second, parrots are popular as pets. Over 70% of all 
extant parrot species are bred in the aviculture industry and kept as pets 
worldwide [@Anderson2003;@Carrete2008;@Drews2001;@Kelly2014;@Li2014;@Su2015;
@Young2012], enabling us to leverage the power of digital video platforms to 
search for evidence of tool use [@Nelson2013]. Third, detailed data on relative
brain sizes [@Flammer2001;@Iwaniuk2005;@Mazengenya2018;@Olkowicz2016;
@Schuck2008;@Silva2017], feeding strategies [@Wilman2016], and shared ancestry
[@Jetz2012] exist for parrots, allowing us to fit the statistical model implied
by Figure \@ref(fig:plotDAG) to the entire parrot order.

We first present the results from our video survey, in which we collated videos
of tool use in parrots from a digital video platform. This survey reveals that a
number of parrots not previously known to use tools are capable of self-care 
tooling. We then map these previously unidentified tool-using species onto the
phylogeny of the parrot order and use a phylogenetic survival cure model to 
(_i_) rank further parrot species that are likely unobserved tool users and 
(_ii_) re-examine key hypotheses regarding the evolutionary drivers and origins
of tool use in parrots.

@@

## Digital video platform reveals self-care tool use in additional parrot species

@@

In total, we found `r nrow(crowdData)` videos of 
`r crowdData %>% pull(IndividualID) %>% unique() %>% length()` individuals from 
`r crowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
parrot species performing behaviours that met the definition of either true tool 
use (`r nrow(trueCrowdData)` videos of 
`r trueCrowdData %>% pull(IndividualID) %>% unique() %>% length()` 
individuals from 
`r trueCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
species) or borderline tool use (`r nrow(bordCrowdData)` videos of 
`r bordCrowdData %>% pull(IndividualID) %>% unique() %>% length()` 
individuals from 
`r bordCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
species). All videos featured pet parrots in captive settings performing
self-care tool use, specifically self-scratching. In 
`r crowdData %>% pull(HumanInfluence) %>% str_starts("No") %>% sum()` of these 
videos, owners did not appear to interact with the subjects. In 
`r crowdData %>% pull(HumanInfluence) %>% str_starts("Yes") %>% sum()` videos, 
there was potential human interaction, either from the owners being in close physical 
contact with the bird (e.g., bird perching on hand), talking to the bird, or 
handing it the tool (which occurred in only two videos). We could not establish 
the degree of human interaction in the remaining 
`r crowdData %>% pull(HumanInfluence) %>% str_starts("Unknown") %>% sum()` 
videos, as sound had been removed or was substituted by music. None of the 
videos featured owners directly rewarding tool use behaviours with food. All 
borderline tool use cases were excluded from further analyses.

Of the `r trueCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
parrot species performing true tool use for self-care, 
`r trueCrowdData %>% filter(Species != "Unknown") %>% group_by(Species, IndividualID) %>% summarise(n = n(), .groups = "drop") %>% group_by(Species) %>% summarise(check = n() > 1, .groups = "drop") %>% pull(check) %>% sum()` 
were represented in our video survey by two or more individuals over multiple 
independent observations. True tool use always involved the subject using an 
object for self-scratching (`r trueCrowdData %>% mutate(BodyPartScratched = str_detect(BodyPartScratched, "Head/Neck")) %>% pull(BodyPartScratched) %>% sum()` 
videos involved scratching the head and/or neck). The most common tool (`r sum(pull(trueCrowdData, EvaluatedTool) == "Feather")`
videos) was a moulted feather. Human-made objects (e.g., pens, spoons, pieces of 
wood, cardboard) were also common. In `r sum(readxl::read_xlsx(tar_read(crowdsourcing_file))$Directionality == "Tool to body", na.rm = TRUE)` 
videos, parrots manipulated the tool while keeping their body still, rather than 
holding the tool still and moving towards it.

According to YouTube video descriptions and owner comments, 
`r trueCrowdData %>% group_by(IndividualID, Sex) %>% summarise(.groups = "drop") %>% pull(Sex) %>% (function(x) sum(x == "Male"))` 
of the individuals performing true tool use for self-care were males and 
`r trueCrowdData %>% group_by(IndividualID, Sex) %>% summarise(.groups = "drop") %>% pull(Sex) %>% (function(x) sum(x == "Female"))` 
were females. No sex information was provided for the remaining 
`r trueCrowdData %>% group_by(IndividualID, Sex) %>% summarise(.groups = "drop") %>% pull(Sex) %>% (function(x) sum(x == "Unknown"))` 
individuals. As owners provided no information on whether sex had been 
established through genetic testing, and sexual dimorphism in parrots is 
rare [@Bercovitz1987;@Hoyo2011], we could not typically ascertain if descriptions 
were accurate. It is unclear if the disproportionately large number of males in 
the sample is a consequence of owners more readily assuming their parrots are 
male when they have not been genetically tested, owners being more likely to own 
or film male parrots, or male parrots exhibiting more self-scratching behaviours 
than female parrots.

Figure \@ref(fig:plotPhylo1) maps the findings from the video survey onto a 
maximum clade credibility phylogeny for the parrot order, plotted alongside 
species previously identified in the scientific literature. Before the video 
survey, 11 parrot species (3%) had been identified as tool users in the 
scientific literature. Across our video survey, we observed true tool use for self-care in 
`r trueCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
species, 5 of which overlapped with the scientific literature and 
`r (trueCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()) - 5` 
of which were novel species. All of the species identified in the video survey 
were cockatoos (_Cacatuidae_), Old World parrots (_Psittacinae_), or neotropical 
parrots (_Arinae_). The most common species in our survey, accounting for 
`r crowdData %>% filter(Species == "Pyrrhura molinae") %>% nrow()` videos from 
`r crowdData %>% filter(Species == "Pyrrhura molinae") %>% pull(IndividualID) %>% unique() %>% length()` 
individuals, was the green-cheeked conure (_Pyrrhura molinae_). In accordance 
with the scientific literature, the video survey found no evidence of tool use 
in any species of Psittaculidae, despite this family containing some of the most 
commonly kept pet species, including lovebirds, lorikeets, and Asian parakeet 
species. Combining both the video survey and the scientific literature, we can 
thus identify 28 total parrot species (7%) capable of true tool use as defined 
by ref [@Shumaker2011], compared to the 11 previously reported.

@@

The identification of self-care tooling behaviours in several new parrot species
in our video survey increases the extent to which phylogeny can explain the 
distribution of a cognitive capacity for tool use in the parrot order. We 
estimated the phylogenetic signal (Pagel's $\lambda$) of tool use using both the
pre-video-survey and post-video-survey data. Pagel's $\lambda$ varies between 0 
and 1, where 0 implies that the distribution of a trait across species is 
unexplained by phylogenetic relatedness and 1 implies that the distribution of a 
trait across species is fully explained by phylogeny. Using the evidence of tool
use from the scientific literature alone (pre-video-survey data; 11 tool-using 
species), we estimated an average posterior Pagel's $\lambda$ of 
`r printnum(mean(tar_read(phySignal1)))` (95% credible interval [`r printnum(quantile(tar_read(phySignal1), 0.025))` 
`r printnum(quantile(tar_read(phySignal1), 0.975))`]; total _n_ = 
`r nrow(tar_read(d))` species). This estimate was moderate-to-strong, but highly 
uncertain. In comparison, combining the evidence from both the literature and 
the video survey (post-video-survey data; 28 tool-using species) resulted in a 
stronger and more certain estimate of phylogenetic signal. With these data, we 
estimated Pagel's $\lambda$ = `r printnum(mean(tar_read(phySignal2)))` (95% CI 
[`r printnum(quantile(tar_read(phySignal2), 0.025))` 
`r printnum(quantile(tar_read(phySignal2), 0.975))`]; total _n_ = 
`r nrow(tar_read(d))` species). Thus, the results of our video survey increase 
the extent to which the distribution of tool use capabilities across parrot 
species can be explained by shared phylogenetic ancestry. This suggests that we
can potentially use phylogenetic information, along with other variables, to 
identify further parrot species that are capable of tool use but have remained 
undetected.

@@

If we combine evidence from both the literature and our video survey, 28 parrot
species in total have demonstrated a capacity for some form of tool use, whether
that be self-care tool use in captivity, tool use in a wild foraging context, or
other tool behaviours that fit the criteria for true tool use [@Shumaker2011].
By analysing these observations alongside data on relative brain size, feeding
strategy, and phylogenetic ancestry, we can (_i_) identify which parrot species
without recorded evidence of tool use are most likely to have a capacity for 
tool use that has as yet gone unobserved, and (_ii_) test theories about the 
evolutionary drivers and origins of technical intelligence in the parrot order.
To this end, we fitted a Bayesian phylogenetic survival cure model to the data
on parrot tool use from both the literature and our video survey.

@@

Our tool use problem has the same features. We are modelling a time-to-event; 
specifically, the amount of "time" (i.e., observation opportunities measured as 
the number of published papers or videos) until the capacity for tool use is 
identified in some form of tool use behaviour. This is right-censored data 
because many species will not have had tool use identified when we measure them.
Moreover, we can assume that a certain proportion of the population will never
experience the event -- that is, they do not have the capacity for tool use and
so we will never observe tool use behaviour no matter how long we measure them
for.

@@

The posterior probabilities shown in Figure \@ref(fig:plotSurvCure1) are 
estimated with uncertainty, so it is difficult to "identify" any particular 
species as having an undetected capacity for tool use. Nevertheless, taking the
sum of all the posterior probabilities for the 
`r nrow(plotSurvCure1$data %>% filter(ToolUse == "Absent"))` species without 
observed evidence of tool use, we can estimate that around 
`r printnum(median(sumProbs), digits = 0)` of those species are likely to be 
undetected tool users (median sum of probabilities = 
`r printnum(median(sumProbs), digits = 2)`, 95% CI
[`r printnum(quantile(sumProbs, 0.025), digits = 2)` 
`r printnum(quantile(sumProbs, 0.975), digits = 2)`]). When combined with the 
species known to use tools, this implies that between 11% and 17% of extant 
parrot species may be tool users.

@@

Since the earliest anecdotes of parrots using tools by Wallace in the 1880s
[@Wallace1869] and more systematic anecdotal reports in the 1970s [@Goodall1971], 
only 11 parrot species (3% of all extant parrots) have been documented as tool 
users in the scientific literature. Our study collated data from a digital video
platform and revealed capabilities for self-care tool use in 17 additional 
parrot species, more than doubling the overall count of tool-using parrots to 
28 species (7%). These species consisted of cockatoos (_Cacatuidae_), Old World
parrots (_Psittacinae_), and neotropical parrots (_Arinae_). Beyond the video 
survey, the strong phylogenetic signal in our dataset allowed us to use 
phylogenetic information, along with other variables, to infer the unobserved
probabilities of tool use across the parrot order. Our phylogenetic survival
cure model incorporated information on phylogenetic history, research effort, 
relative brain size, and feeding specialisation to rank parrot species that were
most likely to be undetected tool users. The sum of probabilities from this 
model implied that between 15 and 41 of the species without evidence of tool use
are likely to have an unobserved capacity for tool use, suggesting that the true
proportion of tool users in the parrot order may be as high as 17%.

All instances of tool use in our video survey met the established criteria for 
tool use in the literature: the manipulation of an unattached object as an 
extension of the animal's body to achieve a goal [@Shumaker2011]. However, a 
striking feature of the dataset is that the 17 additional species identified in
the video survey were exclusively observed using tools for self-scratching.
While parrots in the wild use tools to achieve a variety of goals, including 
extractive foraging and courtship [@Goodman2018;@Heinsohn2017;@OHara2021;
@Wood1984], parrots in captivity rarely ever require tools in an environment 
where food is readily available and there is little or no competition for mates.
Instead, the most commonly observed form of tool use in captive parrots is 
self-scratching. It could be argued that self-scratching is a more cognitively 
simple form of tool use than extractive foraging as it is egocentric, meaning 
that the tool is directed towards oneself rather than an external object. 
Nevertheless, self-scratching meets the definition for a more complex type of 
embodied tool use known as 'tooling', the deliberate generation of a mechanical 
interface by using an object to manipulate another target or surface 
[@Fragaszy2018]. In our dataset, most self-scratching tooling instances involved 
manipulating the tool while keeping the body still, suggesting that these were 
goal-directed, deliberate movements by individuals. Moreover, the fact that we 
found strong phylogenetic clustering of self-scratching tool use from the video 
survey with other examples of tool use from the literature supports a common 
underlying cognitive mechanism. In line with this, some of the species of 
parrots that use tools for self-scratching in captivity also use tools for other 
purposes in the wild [@Auersperg2012;@Bastos2021;@Goodman2018;@OHara2021]. For 
these reasons, we believe that self-care tool use in captivity is no less 
indicative of a cognitive capacity for tool use than other tool behaviours in 
the wild or in captivity. Indeed, if this is not the case, then the field will 
need to re-evaluate current working definitions of true tool use and their 
implications for animal cognition. This justifies our use of self-scratching, 
alongside other wild observations of tool use, to make claims about the 
evolutionary drivers and origins of technical intelligence in the parrot order.

Our data raise the question of why such a wide range of parrot species exhibit 
self-scratching tool use behaviours in captive environments. It is possible that 
these self-care behaviours emerge more often in captivity because pet parrots 
are kept with few or no conspecifics, and therefore must innovate tool use to 
preen inaccessible body areas, such as their heads and necks, which would have 
been otherwise preened by flock members. This may indicate that human 
interaction and allopreening in the absence of a conspecific is insufficient for
pet parrots, as suggested by studies on welfare and chronic stress in lone 
captive parrots [@Aydinonat2014;@Engebretson2006;@Mason2010;@Meehan2003]. Beyond 
social isolation in captivity, other factors that reduce preening efficacy by an 
individual might also lead to the innovation of self-care tooling: for example,
a captive kea housed with conspecifics in a naturalistic environment was shown
to have innovated the use of a pebble tool for self-care likely as a consequence 
of a missing upper mandible [@Bastos2021]. These behaviours might therefore be 
innovated by individuals when ecologically necessary. Critically, however, a 
capacity to innovate a behaviour such as tool use must exist in a species in 
order to manifest in any environment, whether that be captive or in the wild. 
The fact that only a subset of parrot species are innovating self-care tooling 
in response to their captive setting and that self-care tooling is predicted by 
phylogenetic history and encephalisation strongly suggests that self-scratching 
indicates a capacity to produce tool use, which is expressed in different ways 
in captive and wild settings.

Our findings have a number of important implications. First, our findings show 
that current research effort in the scientific literature is insufficient to 
capture the real world occurrence of parrot tool use. If the scientific 
literature had sampled the natural world sufficiently, we would expect to see 
close correspondence between those species reported as tool users in the 
literature and those species the public have filmed performing tool use. 
Instead, we discovered a large discrepancy between these two data sources, both 
in the prevalence of tool use and the species identified, likely due to the 
numerous difficulties inherent to conducting observations in the wild. This 
raises the possibility that other rare behaviours remain yet to be discovered in 
the scientific literature.

@@

@@

@@

In conclusion, we have shown that the scientific literature has insufficiently 
captured the full distribution of tool use in the parrot order. Our 
digital video survey identified novel observations of self-care tool use in 
several parrot species, more than doubling the number of known tool using parrot 
species from 11 to 28. Our phylogenetic model suggested that the true 
proportion of parrot tool users could be as high as 17% of all species in this 
order. These discoveries have implications for theories of the evolutionary 
drivers and origins of tool use in parrots. Beyond parrot tool use, the 
methods used in this study have the potential to be applied to other rarely 
observed behaviours, including tool use in other taxa [@Hunt2013], rhythmic 
entrainment in birds [@Benichov2016;@Patel2009;@tenCate2016;@Wilson2016],
teasing behaviours in primates [@Eckert2020], and tactical deception across all
animals [@BroJorgensen2010;@Byrne1988;@Byrne1991]. We hope that these methods
will continue to uncover a diverse array of ephemeral behaviours that have as
yet gone undetected in the scientific literature.
